Unsupervised embeddings capture latent structure of scientific mobility

### 2020-02-25 paper/current/main.tex
added basic abstract with some references

\section{Introduction}\label{sec:introduction} %{{{

@@

We consider separate variants of $f(r_{ij})$ for geographic distance, $g_{ij}$, and the embedding distance, $d_{ij}$, report the best-fit model of each distance.
For geographic distance, we use the power-law function of the gravity law, with $f(g_{ij})= g^{-\alpha}_{ij}$, where $\alpha$ is estimated from the data. 
For embedding distance , we use the exponential function, with $f(d_{ij}) = e^{-\beta d_{ij}}$, where $\beta$ is estimated from the data. 
To estimate parameters $\alpha$ and $\beta$ are estimated by log-transforming each variant of \ref{eq:gravity_basic}, obtaining Eq.~\ref{eq:linear_flux_geo} and Eq.~\ref{eq:linear_flux_emb} in the form of linear equations, and fitting parameters using linear regression.

@@

%
% Figure - Conditional Probability
%
\begin{figure}
    \centering
    \includegraphics[width=0.8\columnwidth]{\figdir/Math/approxi_proof.pdf}
    \caption{Relationship between the conditional probability $p(j|i)$ from data and approximate functional form Eq. \ref{eq:real_embedding_relationship}}.
    \label{fig:real_embedding_relationship}
\end{figure}

@@

The SGNS model learns an effective distance from trajectories, one which we can visualize the results into two dimensional space by projecting high dimensional organizational vectors into two dimensions vectors using UMAP \cite{mcinnes2018umap} (Fig.~\ref{fig:projection}a).
Overall, organizations are clustered by geography, according to the nations in which they are location;
however, the embedding space also reflects a mix of geographic, historic, cultural, and linguistic relationships between regions at multiple scales.

@@

Six country clusters were identified, reflecting a mixture of the regional, linguistic, historical clusters. 
From top-to-bottom, we identify a primarily East-Asian cluster (orange), a Scandinavian cluster (dark purple), a cluster of British Commonwealth countries (light purple), a central and eastern European cluster (sky blue), a diverse cluster of European and South American clusters (dark green), and finally another diverse cluster consisting of countries bordering the Mediterranean or the Black Sea.
When not defined by geography, clusters appear to be defined by linguistic and historical ties, as is the case for the commonwealth countries. 
By looking closely at the dendrogram and meta-data, there is a tendency that countries are merged as a group if they share the language in the early stage of the deprogram and geographical aspect plays a more major role in the later stage of the dendrogram.

We investigate the extent to which geography and language dictate the formation of clusters using \textit{CluSim}~\autocite{gates2019element}.
\textit{CluSim} is a method that calculates the similarity between the cluster results and, for hierarchical clustering, can adjust the relative importance of membership at different levels of the hierarchy with scaling parameter $r$, acting likes ``zooming lens''.  
If $r$ is high, the similarity is based on the lower levels of the dendrogram , and when $r$ is low, the similarity is based on higher levels of the dendrogram. 
In fig. \ref{fig:clustering}a, we report three cluster similarity curves (continent, language, language family) with varying scaling parameter $r$.
As expected, continent plays a major role at higher levels of clustering process (low $r$), language and language family explain the clustering process more at the lower levels (high $r$). 
This suggests that geography, then language, are important factors contributing to the structure of international scientific mobility.

>>>>

%
% ABSTRACT
%

\begin{abstract}

Scientific mobility, the movement of scientists between organizations, drives innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, jonkers2013return, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige} however the complexity of factors driving global scientific mobility~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career} necessitates simplified models for reasoning about and crafting policy around mobility. 
Common models of mobility, such as the Gravity Model~\autocite{curiel2018citygravity, lewer2008immigrationgravity, jung2008highwaygravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity, simini2012universal}, are a function on the \textit{distance} between places.
However, amid a globalizing scientific enterprise~\autocite{czaika2018globalisation}, geographic distance is only becoming less useful for understanding mobility, and other abstract notions of distance~\autocite{boschma2005proximity} cannot be measured.
Here, we learn an effective distance directly from mobility data using the skip-gram word embedding model~\autocite{mikolov2013word2vec}, treating organizations analogously to words and career trajectories to sentences.
We source publications of disaambiguated authors from the Web of Science database to construct mobility embeddings for authors career trajectories between 2008 and 2019. 
We demonstrate distance between vectors in the learned embedding better explains and predicts mobility between organizations than does geographic distance, a result of an underlying connection between word embeddings and the gravity model of mobility.
In additional to capturing geographic relationships, these embeddings capture latent religious, cultural, social, and linguistic relationships between organizations at multiple scales of granularity. 
Without explicit inclusion of prestige, embeddings also capture complex organizational hierarchies and can be used to re-construct university rankings. 
Our findings demonstrate the potential of learning efficient, dense, and meaningful vector-space representations of scientific mobility directly from the data, one that models a complex notion of inter-organization distance and which points towards a generalized methodology of reasoning about all kinds of mobility.

\end{abstract}

%
% INTRODUCTION
%
\section{Introduction}\label{sec:introduction} %{{{

@@

We consider separate variants of $f(r_{ij})$ for geographic distance, $g_{ij}$, and the embedding distance, $d_{ij}$, report the best-fit model of each distance.
For geographic distance, we use the power-law function of the gravity law, with $f(g_{ij})= g^{-\alpha}_{ij}$, where $\alpha$ is estimated from the data. 
For embedding distance , we use the exponential function, with $f(d_{ij}) = e^{-\beta d_{ij}}$, where $\beta$ is estimated from the data. 
To estimate parameters $\alpha$ and $\beta$ are estimated by log-transforming each variant of \ref{eq:gravity_flux}, obtaining Eq.~\ref{eq:linear_flux_geo} and Eq.~\ref{eq:linear_flux_emb} in the form of linear equations, and fitting parameters using linear regression.

@@

%
% Figure - Conditional Probability
%
\begin{figure}
    \centering
    \includegraphics[width=0.8\columnwidth]{\figdir/Math/approxi_proof.png}
    \caption{Relationship between the conditional probability $p(j|i)$ and approximate functional form Eq. \ref{eq:real_embedding_relationship}}.
    \label{fig:real_embedding_relationship}
\end{figure}

@@

The SGNS model learns an effective distance from trajectories, one which we can visualize by projecting organizational vectors into two dimensions using UMAP \cite{mcinnes2018umap} (Fig.~\ref{fig:projection}a).
Overall, organizations are clustered by geography, according to the nations in which they are location;
however, the embedding space also reflects a mix of geographic, historic, cultural, and linguistic relationships between regions at multiple scales.

@@

Six country clusters were identified, reflecting a mixture of the regional, linguistic, historical clusters. 
From top-to-bottom, we identify a primarily East-Asian cluster (orange), a Scandinavian cluster (dark purple), a cluster of British Commonwealth countries (light purple), a central and eastern European cluster (sky blue), a diverse cluster of European and South American clusters (dark green), and finally another diverse cluster consisting of countries bordering the Mediterranean or the Black Sea.
When not defined by geography, clusters appear to be defined by linguistic and historical ties, as is the case for the commonwealth countries. 
By looking closely at the deprogram and meta-data, there is a tendency that countries are merged as a group if they share the language in the early stage of the deprogram and geographical aspect plays a more major role in the later stage of the dendrogram.

We investigate the extent to which geography and language dictate the formation of clusters using \textit{CluSim}~\autocite{gates2019element}.
\textit{CluSim} is a method that calculates the similarity between the cluster results and, for hierarchical clustering, can adjust the relative importance of membership at different levels of the hierarchy with scaling parameter $r$, acting likes ``zooming lens''.  
If $r$ is high, the similarity is based on the lower levels of the dendrogram, and when $r$ is low, similarity is based on higher levels dendrogram. 
In fig. \ref{fig:clustering}b, we report three cluster similarity curves (continent, language, language family) with varying scaling parameter $r$.
As expected, continent plays a major role at higher levels of clustering process (low $r$), language and language family explain the clustering process more at the lower levels (high $r$). 
This suggests that geography, then language, are important factors contributing to the structure of international scientific mobility.
